### 2021-05-05 thesis/tex/nsiResults.tex
update

\documentclass[draft=True]{revtex4-2}
\newcommand{\zreco}{\ensuremath{\cos{(\theta_z^{reco})}}}
\newcommand{\ztrue}{\ensuremath{\cos{(\theta_z^{true})}}}
\newcommand{\emm}{\ensuremath{\epsilon_{\mu\mu}}}
\newcommand{\emt}{\ensuremath{\epsilon_{\mu\tau}}}
\newcommand{\eet}{\epsilon_{e\tau}}
\newcommand{\eem}{\epsilon_{e\mu}}
\newcommand{\ett}{\ensuremath{\epsilon_{\tau\tau}}}
\newcommand{\ep}{\ensuremath{\epsilon^\prime}}
\renewcommand{\ne}{\nu_e}
\newcommand{\Ereco}{E^{reco}}
\newcommand{\Etrue}{E^{true}}
\newcommand{\nm}{\nu_\mu}
\newcommand{\nt}{\nu_\tau}
\newcommand{\ane}{\bar\nu_e}
\newcommand{\anm}{\bar\nu_\mu}
\newcommand{\ant}{\bar\nu_\tau}
\newcommand{\dm}{\Delta m^2_{31}}
\newcommand{\sth}{\sin^2(2\theta_{23})}

\usepackage{physics,amsmath, amsfonts, siunitx, amssymb, graphicx}
\usepackage[utf8]{inputenc}
\usepackage{hyperref}
\begin{document}
\section{Methodology}
The neutrino flux at the detector is calculated by propagating the atmospheric neutrino flux~\cite{hondapaper} through the Earth by solving the 
Schrödinger equation for varying density. The Earth density profile is taken from the PREM~\cite{PREM}. The baseline for a given trajectory is determined using an average neutrino
production height of 15 km and an Earth radius of 6371 km. %TODO: add more here?

The oscillation probability $P_{\alpha \beta}$ acts as a weight to the atmospheric flux, yielding the propagated flux at detector level for flavor $\beta$ as 
\begin{align}\label{eq:propFlux}
    \phi_\beta^\text{det} = \sum_\alpha P_{\alpha\beta} \phi_\alpha^\text{atm} \,,
\end{align}
where we sum over the initial lepton flavors $\alpha \in \{e,\mu, \bar{e}, \bar{\mu}\}$. To illustrate the impact of $\emt$ on both probability
and flux level, we plot the oscillograms resulting from Eq.\ref{eq:propFlux} in Fig.~\ref{fig:flux_ratio}. In the left panel, we have marked the region in which 99\% of the 
DeepCore cascade events originating from $\ne$ and $\nt$ fluxes are contained. In the right panel, we show the two regions in which 99\% of the IceCube 
and DeepCore track events originating from $\nm$ fluxes are contained. Starting with the $\nm$ flux ratio, we see that the only clear signal discernible to the IceCube detector
is a flux deficiency of a factor of $\sim 10^2$ from core-crossing neutrinos within a zenith range of $\ztrue > -0.87$. DeepCore, on the other hand, 
observes multiple fringes of flux surpluses with a factor $\sim 10^1$. The strongest surplus at \SI{20}{\GeV} is very weakly zenith dependent, a stark contrast to the
energy-independent but zenith-sensetive IceCube deficiency.

For the fluxes which drives cascades, we have to resort to the DeepCore detector alone. %TODO: fill more
Here we see a somewhat weaker signal, this time a zenith-independent deficiency. 
%TODO: fill with conclusion

\begin{figure}[!tb]
   \begin{center}
      \includegraphics[width=0.4\linewidth]{figures/cascade_flux_ratio.pdf}
      \hspace{1.5cm}
      \includegraphics[width=0.4\linewidth]{figures/track_flux_ratio.pdf}
   \end{center}
   \caption{Ratio of propagated NSI to SI atmospheric fluxes at detector level. We set $\emt = -0.05$, and all other NSI parameters to zero. 
   Dotted (dashed) lines show the regions in which 99\% of the 
   DeepCore (IceCube) MC events are contained. \emph{Left panel:} Propagated fluxes of $\ne$ and $\nt$ neutrinos and anti-neutrinos. 
   \emph{Right panel:} Propagated fluxes of $\nm$ neutrinos and anti-neutrinos}\label{fig:flux_ratio}%TODO: normalize color scale and fix limits
\end{figure}

\subsection{IceCube}\label{ch:ICmethod}
The event rate for each bin reads
\begin{align}\label{eq:ICevents}
    N_{ij} &= T \int_{(\cos{\theta_z^r})_i}^{(\cos{\theta_z^r})_{i+1}} \dd \cos{\theta^r_z} \int_{E^r_{j}}^{E^r_{j+1}} \dd E^r \int_0^\pi R(\theta^r,\theta^t) \dd \cos{\theta^t} \int_0^\infty R(E^r,E^t) \dd E^t
    \times \left[ \sum_\beta \phi_\beta^\text{det}  A^\text{eff}_\beta\right]\,,
\end{align}
where $T$ is the live time of the detector, and $A^\text{eff}_\beta$ its effective area for flavor $\beta$. We use the effective area of the 86 string configuration made public by the IceCube collaboration~\cite{ICaeff}. $R(x^r,x^t)$ is a resolution function, 
which is responsible for the smearing between the reconstructed and true parameters $x^r$ and $x^t$, respectively. We assume a log-normal distribution, giving it the form 
\begin{align}
    R(x^r, x^t) = \frac{1}{\sqrt{2\pi} \sigma_{x^r}x^r} \exp\left[-\frac{(\log x^r-\mu(x^t))^2}{2\sigma_{x^r}^2}\right]\,.
\end{align}
As seen in Fig.~\ref{fig:IC_MC_counts}, the energy reconstruction is biased. To model this relationship between $\Etrue$ and $\Ereco$, we train a Gaussian process regressor on the dataset~\cite{IC2016}, from which
we can extract a predicted mean and standard deviation for a given $E^{reco}$. We then take the $\Etrue$ points of the 99th percentile of each distribution to obtain
the limits of $\Etrue$ at which to integrate over. We have no angular resolution function since the angle resolution in Icecube for track-like events is less than $\SI{2}{\degree}$, making $\ztrue$ coincide with $\zreco$ for our study~\cite{IC2020}. 
The data is from the IC86 sterile analysis~\cite{IC2020}.
\begin{figure}[!tb]
    \begin{center}
       \includegraphics[width=0.4\linewidth]{figures/IC_MC_gpr.pdf} %TODO: another fig here? Aeff?
    \end{center}
    \caption{Relationship between the true and reconstructed muon energy in the IceCube MC sample~\cite{IC2016}}\label{fig:IC_MC_counts}. Shaded area shows the $99.9$th percentile limits predicted by the regressor trained on this set.
 \end{figure}

\subsection{DeepCore}\label{ch:DCmethod}
In this part, we use the publically available DeepCore data sample~\cite{DC2019data} which is an updated version of what was used by the 
IceCube collaboration in a $\nu_\mu$ disapprearance analysis~\cite{DC2018mudisappearance}.

The detector systematics include ice absorption and scattering, and overall, lateral, and head-on optical efficiencies of the DOMs. 
They are applied as correction factors using the best-fit points from the DeepCore 2019 $\nu_\tau$ appearance 
analysis~\cite{DC2019tauappearance}.

The data include 14901 track-like events and 26001 cascade-like events, both divided into eight 
$ \log_{10}E^{reco} \in [0.75,1.75]$ bins, and eight $\zreco \in [-1,1]$ bins. Each event has a Monte Carlo weight $w_{ijk,\beta}$,
from which we can construct the event count as
\begin{align}\label{eq:MCevents}
    N_{ijk} &= C_{ijk}\sum_{\beta}w_{ijk,\beta} \phi_\beta^\text{det}\,,
\end{align}
where $C_{k\beta}$ is the correction factor from the detector systematic uncertainty and $\phi_\beta^\text{det}$ is defined as Eq.~\ref{eq:propFlux}. We have now substituted the effect of the Gaussian smearing 
by treating the reconstructed and true quantities as a migration matrix.

The oscillation parameters used on our DeepCore simulations are from the
best-fit in the global analysis in~\cite{nufit}: $\theta_{12} = \SI{33.44}{\degree},\, \theta_{13} = \SI{8.57}{\degree},\, \Delta m^2_{21} =  \SI{7.42}{\electronvolt^2}$, and we 
marginalize over $\dm$ and $\theta_{23}$.

We plot the event pull $(N_{NSI} - N_{SI})/\sqrt{N_{SI}}$ where $N_{(N)SI}$ are the numbers of expected events
assuming (non-)standard interactions in Fig.~\ref{fig:event_pulls}. This gives the normalized difference in the
number of expected events at the detector, and illustrates the expected sensitivity of DeepCore for the NSI parameters.

\begin{figure}[!tb]
   \begin{center}
      \includegraphics[width=0.7\linewidth]{figures/DC_event_pulls.pdf}
      \includegraphics[width=0.7\linewidth]{figures/PINGU_event_pulls.pdf}
   \end{center}
   \caption{Expected pulls of predicted event numbers for DeepCore and PINGU after 3 years. We compare the NSI event count with $\emt=-0.05$
    to the standard interaction count}\label{fig:event_pulls} %TODO: remake these with clearer titles and better labels
\end{figure}

\subsection{PINGU}\label{ch:PINGUmethod}
The methodology behind the PINGU simulations are the same as with our DeepCore study~\ref{ch:DCmethod}. We use the public MC~\cite{PINGUdata}, which allows us to construct the event count as in Eq.~\ref{eq:MCevents}.
However, since no detector systematics is yet modelled for PINGU, the correction factors $C_{ijk}$ are all unity.
As with the DeepCore data, the PINGU MC is divided into eight 
$\log_{10}E^{reco} \in [0.75,1.75]$ bins, and eight $\zreco \in [-1,1]$ bins for both track- and cascade-like events. 
We generate "data" by predicting the event rates at PINGU with the following best-fit parameters from~\cite{nufit}, except for the CP-violating phase which is set to zero for simplicity.

\begin{align}\label{eq:PINGUparams}
    &\Delta m^2_{21} =  \SI{7.42e-5}{\electronvolt^2},\hspace{0.5em} \dm =  \SI{2.517e-3}{\electronvolt^2}, \nonumber \\
    &\theta_{12} = \SI{33.44}{\degree},\hspace{1em} \theta_{13} = \SI{8.57}{\degree},\hspace{1em} \theta_{23} = \SI{49.2}{\degree}, \hspace{1em} \delta_\text{CP} = 0\,.
\end{align}

\section{Results}
For our analyses, we define our $\chi^2$ as
\begin{align} \label{eq:ICchisq}
    \chi^{2}(\hat{\theta},\alpha,\beta, \kappa)=\sum_{ijk} \frac{\left(N^\text{th}-N^\text{data}\right)_{ijk}^{2}}
    {\left(\sigma^\text{data}_{ijk}\right)^{2} + \left(\sigma^\text{syst}_{ijk}\right)^{2}}+ 
    \frac{(1-\alpha)^2}{\sigma_\alpha^2} + \frac{\beta^2}{\sigma_\beta^2}\,
\end{align}
where we minimize over the model parameters $\hat{\theta} \in \{\dm, \theta_{23}, \ep, \emt\}$, the penalty terms $\alpha$ and $\beta$, and the free parameter $\kappa$.
$N_{ijk}^\text{th}$ is the expected number of events from theory, and $N_{ijk}^\text{data}$ is the observed number of events in that bin. We set $\sigma_\alpha = 0.25$ as the atmospheric flux normalization error, and $\sigma_\beta = 0.04$ as the zenith angle slope error~\cite{hondapaper}. 
The observed event number has an associated Poissonian uncertainty $\sigma_{ijk}^\text{data} = \sqrt{N_{ijk}^\text{data}}$.

@@

For DeepCore and PINGU, the event count takes the form
\begin{align}
    N^\text{th}_{ijk} = \alpha\left[1+\beta \zreco_i \right] N_{ijk}(\hat{\theta}) + \kappa N_{ijk}^{\mu_{atm}}\,,
\end{align}
with $N_{ijk}(\hat{\theta})$ from Eq.~\ref{eq:MCevents}. $N_{ijk}^{\mu_{atm}}$ is the muon background, which is left to float freely in the DeepCore analysis.
The background at PINGU can be considered neglible to first order~\cite{PINGUdata}, and we thus put $\kappa=0$ when calculating the PINGU $\chi^2$\,.
For IceCube, we set $\sigma_{ijk}^\text{syst} = f\sqrt{N_{ijk}^\text{data}}$.
For DeepCore, we use the provided systematic error distribution which accounts for both uncertanties in the finite MC statistics and in the data-driven 
muon background estimate~\cite{DC2019data}.

First, we set all standard oscillation parameters to their current best-fit values of Eq.~\ref{eq:PINGUparams}, except for $\dm$ and $\theta_{23}$, 
which we marginalize over their $3\sigma$ ranges of \SI{2.435e-3} to \SI{2.598e-3}{\electronvolt^2} and \SI{40.1} to \SI{51.7}{\degree} respectively. %TODO: make unit ranges nicer.

 \begin{figure}[!tb]
    \begin{center}
       \includegraphics[width=0.45\linewidth]{figures/PINGU_3D_ett.pdf}
       \includegraphics[width=0.45\linewidth]{figures/PINGU_3D_emt.pdf}
       \includegraphics[width=0.45\linewidth]{figures/PINGU_3D_eem.pdf}
       \includegraphics[width=0.45\linewidth]{figures/PINGU_3D_eet.pdf}
    \end{center}
    \caption{The expected values of $\Delta \chi^2$ after three years of PINGU data. $\dm$ and $\theta_{23}$ and have been marginalized out, and all NSI 
    parameters not shown in each plot are set to zero.}\label{fig:PINGU_3D}
 \end{figure}

 \begin{figure}[!tb]
    \begin{center}
       \includegraphics[width=0.4\linewidth]{figures/DC_emm.pdf}
       \includegraphics[width=0.4\linewidth]{figures/DC_emt.pdf}
    \end{center}
    \caption{The expected values of $\Delta \chi^2$ after three years of PINGU data. $\dm$ and $\theta_{23}$ and have been marginalized out, and all other NSI 
    parameters are set to zero. The black lines show the 90\% credibility region from~\cite{demidov}}\label{fig:DC_3D_et}
 \end{figure}

For the joint analysis, we follow the parameter goodness-of-fit prescription~\cite{maltoni2003} and construct the joint $\chi^2$ as 
\begin{align}
    \chi^2_\text{joint} = \chi^2_\text{IC} + \chi^2_\text{DC} + \chi^2_\text{P} - \chi^2_{IC,min} - \chi^2_\text{DC,min} - \chi^2_\text{P,min}\,
\end{align}
with test statistic $\chi^2_\text{joint,min}$.

\begin{figure}[!tb]
    \begin{center}
       \includegraphics[width=0.4\linewidth]{figures/emm_deltachi.pdf}
    \end{center}
    \caption{}\label{fig:emm_deltachi}
 \end{figure}

 \begin{figure}[!tb]
    \begin{center}
       \includegraphics[width=0.4\linewidth]{figures/emm_emp_contour.pdf}
    \end{center}
    \caption{}\label{fig:emm_emp_contour}
 \end{figure}
%At $90\%$ CL:
%PINGU: $-0.046 < \emm < 0.041$
%DC: $-0.045 < \emm < 0.041$
%joint: $-0.038 < \emm < 0.033$

%TODO: table with individual and combined chisq values. Discussion on impact of IC on chisq.

 \newpage
\begin{tabular}{p{55mm}p{55mm}p{55mm}}
   DeepCore (2017)
      \begin{itemize}
         \item[$\checkmark$] Honda atmospheric fluxes
         \item[$\times$] Only look at tracks and $\emt$
         \vspace{1em}  
         \item[$\times$] DC Monte Carlo from an older dataset 
         \item[$\times$] 8 E bins from $\SI{6.3}{\electronvolt^2}$ to $\SI{56}{\electronvolt^2}$
         \item[$\times$] 8 z bins from -1 to 0 
         \item[$\times$] Use "Overall" and "relative $\ne$ to $\nm$" normalization
         \item[$\times$] Prior on spectral index
         \item[$\times$] No zenith angle normalization
         \item[$\checkmark$] No priors on $\dm, \theta_{23},\theta_{13}$
      \end{itemize} &
    Demidov (2020) DC analysis
      \begin{itemize}
         \item[$\checkmark$] Honda atmospheric fluxes
         \item[$\checkmark$] Looks at tracks + cascades for $\emt$ and $\ett$
         \item[$\checkmark$] Data and Monte Carlo from DC 2018
         \item[$\checkmark$] 8 E bins from $\SI{5.6}{\electronvolt^2}   $ to $\SI{56}{\electronvolt^2}$
         \item[$\checkmark$] 8 z bins from -1 to 1
         \item[$\times$] Use "Overall" and "relative $\ne$ to $\nm$" normalization
         \item[$\times$] Prior on spectral index
         \item[$\times$] No zenith angle normalization
         \item[$\checkmark$] No priors on $\dm, \theta_{23}$
         \item[$\checkmark$] Fixes $\Delta m^2_{21}, \theta_{12}, \theta_{13}$
         \item[$\times$] Uncertainty on hadron production in atmosphere
         \item[$\times$] Uncertainty on neutrino nucleon cross section 
      \end{itemize} &
    This DC+PINGU analysis
      \begin{itemize}
         \item[$\checkmark$] Honda atmospheric fluxes
         \item[$\checkmark$] Tracks and cascades for all flavors
         \vspace{1em} 
         \item[$\checkmark$] Reco $\to$ true mapping from Monte Carlo migration matrix
         \item[$\checkmark$] 8 E bins from $\SI{5.6}{\electronvolt^2}$ to $\SI{56}{\electronvolt^2}$
         \item[$\checkmark$] 8 zenith angle bins from -1 to 1
         \item[$\checkmark$] Flux normalization uncertainty of 25\%
         \item[$\checkmark$] Zenith angle uncertainty of 4\% 
         \item[$\checkmark$] No priors on oscillation parameters 
         \item[$\checkmark$] Marginalize $\dm$ and $\theta_{23}$. All other oscillation parameters are fixed.
      \end{itemize} 
\end{tabular}

\begin{table}
   \begin{center}
   \begin{tabular}{lcccc}
           \hline \hline & \multicolumn{3}{c} {\text {Best fit}}& \\
           \cline { 2 - 4 } Parameter & $\dm$ & $\theta_{23}$  & $\epsilon$ & 90\% CL of $\epsilon$ \\
           \hline
           $\ett$ &-& - & - & -\\
           $\emt$ &0.002435& 0.76736 & -0.005 & -0.015, 0.005\\
           $\eem$ &-& - & - & -\\
           $\eet$ &0.002435 & 0.7673630944601736 & 0.05 & -0.072, 0.138\\
           \hline
       \label{table:results}
   \end{tabular}
   \end{center}
   \caption{DeepCore results}
\end{table}

\begin{table}
   \begin{center}
   \begin{tabular}{lcccc}
           \hline \hline & \multicolumn{3}{c} {\text {Best fit}}& \\
           \cline { 2 - 4 } Parameter & $\dm$ & $\theta_{23}$  & $\epsilon$ & 90\% CL of $\epsilon$ \\
           \hline
           $\ett$ &-& - & - & -\\
           $\emt$ &-& - & - & -\\
           $\eem$ &-& - & - & -\\
           $\eet$ &0.002517 & 0.8587019919812102 & 0.05 & -0.105, 0.101\\
           \hline
       \label{table:results}
   \end{tabular}
   \end{center}
   \caption{DeepCore results}
\end{table}

\bibliographystyle{elsarticle-num}
\bibliography{article}
\end{document}

>>>>

\documentclass[draft=True]{thesis}
\usepackage[margin=0.5in]{geometry}
\usepackage{graphicx, slashed, siunitx}
\usepackage[utf8]{inputenc}
\usepackage{xr-hyper}
\input{commands}
\begin{document}
\section{$\chi^2$ minimization}
For our analyses, we define our $\chi^2$ as
\begin{align} \label{eq:ICchisq}
    \chi^{2}(\hat{\theta},\alpha,\beta, \kappa)=\sum_{ijk} \frac{\left(N^\text{th}-N^\text{data}\right)_{ijk}^{2}}
    {\left(\sigma^\text{data}_{ijk}\right)^{2} + \left(\sigma^\text{syst}_{ijk}\right)^{2}}+ 
    \frac{(1-\alpha)^2}{\sigma_\alpha^2} + \frac{\beta^2}{\sigma_\beta^2}\,
\end{align}
where we minimize over the model parameters $\hat{\theta} \in \{\dm, \theta_{23}, \ep, \emt\}$, the penalty terms $\alpha$ and $\beta$, and the free parameter $\kappa$.
$N_{ijk}^\text{th}$ is the expected number of events from theory, and $N_{ijk}^\text{data}$ is the observed number of events in that bin. We set $\sigma_\alpha = 0.25$ as the atmospheric flux normalization error, and $\sigma_\beta = 0.04$ as the zenith angle slope error~\cite{hondapaper}. 
The observed event number has an associated Poissonian uncertainty $\sigma_{ijk}^\text{data} = \sqrt{N_{ijk}^\text{data}}$.

@@

For DeepCore and PINGU, the event count takes the form
\begin{align}
    N^\text{th}_{ijk} = \alpha\left[1+\beta \zreco_i \right] N_{ijk}(\hat{\theta}) + \kappa N_{ijk}^{\mu_{atm}}\,,
\end{align}
with $N_{ijk}(\hat{\theta})$ from Eq.~\ref{eq:MCevents}. $N_{ijk}^{\mu_{atm}}$ is the muon background, which is left to float freely in the DeepCore analysis.
The background at PINGU can be considered neglible to first order~\cite{PINGUdata}, and we thus put $\kappa=0$ when calculating the PINGU $\chi^2$\,.
For IceCube, we set $\sigma_{ijk}^\text{syst} = f\sqrt{N_{ijk}^\text{data}}$.
For DeepCore, we use the provided systematic error distribution which accounts for both uncertanties in the finite MC statistics and in the data-driven 
muon background estimate~\cite{DC2019data}.
\end{document}
